\documentclass{article}
\begin{document}

Notes on: \citet{Cole:bioRxiv-2020.06.01.127555}. Ancient Admixture
into Africa from the ancestors of non-Africans.

``Here we present SMCSMC, a Bayesian method for inference of
time-varying population sizes and directional migration rates
under the coalescent-with-recombination model, to study ancient
demographic events. We find evidence for substantial migration
from the ancestors of present-day Eurasians into African groups
between 40 and 70 thousand years ago, predating the divergence of
Eastern and Western Eurasian lineages. This event accounts for
previously unexplained genetic diversity in African populations,
and supports the existence of novel population substructure in the
Late Middle Paleolithic.''

``he forward and backward direction (Fig. 2); full results are given
in Supplemental Section S5. We find that SMCSMC has good power to
detect backward migration pulses up to 60kya (median ratio of inferred
and true IMF, 0.91), while power drops off at 70kya (IMF ratio
0.46). In the pure backward migration case, some forward migration is
falsely inferred, but this is always substantially less than the
inferred backward migration (median ratio inferred forward to true
backward IMF, 0.37; true migration peak 60kya). However, in the case
of true forward migration as well as bidirectional migration, roughly
equal mixtures of forward and backward migration are inferred
(Fig. 2). We conclude that in the epoch 40–70kya the forward and
bidirectional scenarios are difficult to distinguish from each other,
but both can be distinguished from backward migration, the only
scenario resulting in substantially different inferred backward and
forward migration.''

``SMCSMC infers substantially lower African Ne than MSMC in the period
80kya–300kya. In addition, while MSMC inferences show convergence of
African and Eurasian ancestral Ne estimates only around 300kya,
inferences from SMCSMC indicate convergence at 150kya (Fig. 3a),
closer to the hypothesized time of the diversification of the
ancestral lineages prior to the main out-of-Africa migration episode''

``we found that for simulated African populations that had received
substantial migration from the simulated Eurasian population either
through backward or bidirectional migration, inferred Ne values indeed
were substantially inflated compared to true values (Fig. 3c,d), while
this effect was not seen when forward (African-to-Eurasian) migration
was simulated (Fig. 3e). Similarly, single-population Eurasian Ne
estimates were inflated in the presence of forward and bidirectional
migration, but not backward migration (Supplemental Figs. S8–S10). In
contrast, when using a model that includes migration, inferred African
Ne do not show inflation in any of the three scenarios (Fig. 3c-e). We
conclude that the inferences from SMCSMC and MSMC are compatible with
substantial back-migration from ancestral Eurasians into Africans, but
not substantial bidirectional or forward migration.''

Editorial comment: ``when using a model'' should be ``when using
SMCSMC under a model''.

``In addition to branches in local genealogical trees, SCRM retains
non-local branches in the ancestral recombination graph (ARG) within a
user-specified sliding window. In the limit of a chromosome-sized
windows SCRM is equivalent to the coalescent with recombination, while
for a zero-length window it is equivalent to the sequentially
Markovian coalescent (SMC’) [6, 7]; we use a 100kb sliding window to
approximate the CwR and improve accuracy over SMC’ while retaining
tractable inference.''

\end{document}
